---
title: "Untitled"
author: "Xinyue_Cui"
date: "2023-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(formattable)
library(RColorBrewer)
MB_Xeno <- read_csv("/Users/cuixinyue/Desktop/MIBI/MB_Xeno/MIBI-MB-Xeno-Pilot_scdata (1).csv")
```

#cluster_id count
```{r}
Cluster_count_all <- (
  dplyr:: count(MB_Xeno, cluster_id) 
  |> dplyr::rename(count = n) 
  |> dplyr::mutate(cluster_id = forcats::fct_reorder(cluster_id, count))
  #|> dplyr::arrange(desc(count)) 
  |> ggplot2::ggplot(aes(x = cluster_id, y = count)) 
  + geom_bar(stat = "identity", width = 0.7
             ,fill="#f68060") 
             #alpha=.6, width=.5)
  + coord_flip()
  + xlab("Cluster_id") + ylab("Count")
  + theme(text = element_text(size = 18)) 
  + ggtitle("")
  + theme_bw()
)
Cluster_count_all
```
# unique cluster_id count in each mouse
```{r}
mouse_cluster <- (
  MB_Xeno
  |> dplyr::group_by(mouse_id)
  |> dplyr::summarise(unique_clusterid_Count = length(unique(cluster_id)))
  |> dplyr::mutate(mouse_id = factor(mouse_id))
  #|> dplyr::mutate(sample_id = factor(paste0("Sample_", sample_id)))
  #|> dplyr::mutate(sample_id = forcats::fct_reorder(sample_id, unique_mmCount))
  |> ggplot2::ggplot(aes(x = mouse_id, y = unique_clusterid_Count))
  + geom_bar(stat = "identity", width = 0.7
             ,fill="#f68060") 
             #alpha=.8, width=.7)
  #+ coord_flip()
  + xlab("Mouse Id") + ylab("Unique Cluster id Count")
  + ggtitle("")
  + theme(text = element_text(size = 18)) 
  + scale_x_discrete(guide = guide_axis(angle = 90))
  + theme_bw()
)
sample_cluster
```

```{r}
arrange <- ggpubr::ggarrange(
  Cluster_count_all, mouse_cluster,
   labels = c("A", "B"),
   widths = 10, heights = 6) +
  theme(text = element_text(size = 18)) 

# ggsave(filename = "sample_cellType_plots.eps", plot = arrange, device = "eps",
#        width = 10, height = 5,
#        path = here::here("Output", "Figures"))
```

# Phenotype-Sample distribution
```{r}
# number of cells in each unique combination of sample id and cluster id
cluster_count <- dplyr::count(MB_Xeno, sample_id, cluster_id) %>% 
  dplyr::rename(count = n)
head(cluster_count)
```
```{r}
# histogram 1(all count)
 hist(cluster_count$count,
     main = "Histogram of all clusterid count",
     xlab = "Clusterid count", ylab = "Counts", labels = cluster_count$count) 

# histogram 2(count up to 10000)
hist(cluster_count$count[cluster_count$count < 10000],
     main = "Histogram of cell type count < 10000",
     xlab = "Cluster type count", ylab = "Counts",
     breaks = 20)
# histogram 2(count up to 1000)
hist(cluster_count$count[cluster_count$count < 1000],
     main = "Histogram of cell type count < 1000",
     xlab = "Cluster type count", ylab = "Counts",
     breaks = 30)
# histogram 2(count up to 250)
hist(cluster_count$count[cluster_count$count < 250],
     main = "Histogram of cell type count < 250",
     xlab = "Cluster type count", ylab = "Counts",
     breaks = 20)

# histogram 2(count up to 100)
hist(cluster_count$count[cluster_count$count < 100],
     main = "Histogram of cell type count < 100",
     xlab = "Cluster type count", ylab = "Counts",
     breaks = 20)
```
## split this dataset into 7 segments based on different unique values of mouse_id. 
```{r}
segments <- split(MB_Xeno, MB_Xeno$mouse_id)
MB_Xeno_c1m1 <- segments$c1m1
length(unique(MB_Xeno_c1m1$sample_id))
MB_Xeno_c1m4 <- segments$c1m4
MB_Xeno_c2m1 <- segments$c2m1
MB_Xeno_c2m2 <- segments$c2m2
MB_Xeno_c3m2 <- segments$c3m2
MB_Xeno_c3m5 <- segments$c3m5
MB_Xeno_c5m1 <- segments$c5m1
```

# phenotype-sample distribution heatmap for c1m1
```{r}
# phenotype-sample distribution heatmap
 p_c1m1 <-MB_Xeno_c1m1 %>% 
   with(table(sample_id, cluster_id)) %>% 
  as.data.frame() %>% 
  mutate(sample_id = paste("Sample", sample_id)) %>%
  mutate(count = cut(as.numeric(Freq), breaks = c(-1, 1.1, 5.1, 10.1, 
                                   25.1, 50.1, 100.1, 250.1, 500.1, 1000.1, 2000.1, max(Freq, na.rm = TRUE)+1),
                     labels = c("0-1", "1-5", "5-10", "10-25", "25-50",
                                "50-100", "100-250", "250-500", "500-1000", "1000-2000", ">2000"))) %>% 
  mutate(count = factor(as.character(count), levels = rev(levels(count)))) %>% 
  
 ggplot(mapping = aes(x = cluster_id, y = sample_id, fill = count)) +
    geom_tile(colour = "white", linewidth = 0.3) +
  
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Sample"))+
    labs(x = "Phenotype", y = " sample_id", title = "Phenotype-Sample Distribution for mouse_id = c1m1")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +
    
    # color palette
    #scale_fill_manual(values = rev(brewer.pal(11, "RdGy")), na.value = "azure4") + #RdGy RdBu
    scale_fill_manual(values = brewer.pal(11, "RdBu"), na.value = "azure4") + 
  
    # theme
    theme_grey(base_size = 7)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.2, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     )+
    #scale_x_discrete(guide = guide_axis(n.dodge = 2)) # separate x label into two levels
    scale_x_discrete(guide = guide_axis(angle = 45))

```
# phenotype-sample distribution heatmap for c1m4
```{r}
p_c1m4 <- MB_Xeno_c1m4%>% 
   with(table(sample_id, cluster_id)) %>% 
  as.data.frame() %>% 
  mutate(sample_id = paste("Sample", sample_id)) %>%
  mutate(count = cut(as.numeric(Freq), breaks = c(-1, 1.1, 5.1, 10.1, 
                                   25.1, 50.1, 100.1, 250.1, 500.1, 1000.1, 2000.1, max(Freq, na.rm = TRUE)+1),
                     labels = c("0-1", "1-5", "5-10", "10-25", "25-50",
                                "50-100", "100-250", "250-500", "500-1000", "1000-2000", ">2000"))) %>% 
  mutate(count = factor(as.character(count), levels = rev(levels(count)))) %>% 
  
  ggplot(mapping = aes(x = cluster_id, y = sample_id, fill = count)) +
    geom_tile(colour = "white", linewidth = 0.3) +
  
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Sample"))+
    labs(x = "Phenotype", y = " sample_id", title = "Phenotype-Sample Distribution for mouse_id = c1m4")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +
    
    # color palette
    #scale_fill_manual(values = rev(brewer.pal(11, "RdGy")), na.value = "azure4") + #RdGy RdBu
    scale_fill_manual(values = brewer.pal(11, "RdBu"), na.value = "azure4") + 
  
    # theme
    theme_grey(base_size = 7)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.2, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     )+
    #scale_x_discrete(guide = guide_axis(n.dodge = 2)) # separate x label into two levels
    scale_x_discrete(guide = guide_axis(angle = 45))

```
# phenotype-sample distribution heatmap for c2m1
```{r}
p_c2m1 <- MB_Xeno_c2m1%>% 
   with(table(sample_id, cluster_id)) %>% 
  as.data.frame() %>% 
  mutate(sample_id = paste("Sample", sample_id)) %>%
  mutate(count = cut(as.numeric(Freq), breaks = c(-1, 1.1, 5.1, 10.1, 
                                   25.1, 50.1, 100.1, 250.1, 500.1, 1000.1, 2000.1, max(Freq, na.rm = TRUE)+1),
                     labels = c("0-1", "1-5", "5-10", "10-25", "25-50",
                                "50-100", "100-250", "250-500", "500-1000", "1000-2000", ">2000"))) %>% 
  mutate(count = factor(as.character(count), levels = rev(levels(count)))) %>% 
  
  ggplot(mapping = aes(x = cluster_id, y = sample_id, fill = count)) +
    geom_tile(colour = "white", linewidth = 0.3) +
  
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Sample"))+
    labs(x = "Phenotype", y = " sample_id", title = "Phenotype-Sample Distribution for mouse_id = c2m1")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +
    
    # color palette
    #scale_fill_manual(values = rev(brewer.pal(11, "RdGy")), na.value = "azure4") + #RdGy RdBu
    scale_fill_manual(values = brewer.pal(11, "RdBu"), na.value = "azure4") + 
  
    # theme
    theme_grey(base_size = 7)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.2, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     )+
    #scale_x_discrete(guide = guide_axis(n.dodge = 2)) # separate x label into two levels
    scale_x_discrete(guide = guide_axis(angle = 45))

```
# phenotype-sample distribution heatmap for c2m2
```{r}
p_c2m2 <- MB_Xeno_c2m2%>% 
   with(table(sample_id, cluster_id)) %>% 
  as.data.frame() %>% 
  mutate(sample_id = paste("Sample", sample_id)) %>%
  mutate(count = cut(as.numeric(Freq), breaks = c(-1, 1.1, 5.1, 10.1, 
                                   25.1, 50.1, 100.1, 250.1, 500.1, 1000.1, 2000.1, max(Freq, na.rm = TRUE)+1),
                     labels = c("0-1", "1-5", "5-10", "10-25", "25-50",
                                "50-100", "100-250", "250-500", "500-1000", "1000-2000", ">2000"))) %>% 
  mutate(count = factor(as.character(count), levels = rev(levels(count)))) %>% 
  
  ggplot(mapping = aes(x = cluster_id, y = sample_id, fill = count)) +
    geom_tile(colour = "white", linewidth = 0.3) +
  
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Sample"))+
    labs(x = "Phenotype", y = " sample_id", title = "Phenotype-Sample Distribution for mouse_id = c2m2")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +
    
    # color palette
    #scale_fill_manual(values = rev(brewer.pal(11, "RdGy")), na.value = "azure4") + #RdGy RdBu
    scale_fill_manual(values = brewer.pal(11, "RdBu"), na.value = "azure4") + 
  
    # theme
    theme_grey(base_size = 7)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.2, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     )+
    #scale_x_discrete(guide = guide_axis(n.dodge = 2)) # separate x label into two levels
    scale_x_discrete(guide = guide_axis(angle = 45))

```
# phenotype-sample distribution heatmap for c3m2
```{r}
p_c3m2 <- MB_Xeno_c3m2%>% 
   with(table(sample_id, cluster_id)) %>% 
  as.data.frame() %>% 
  mutate(sample_id = paste("Sample", sample_id)) %>%
  mutate(count = cut(as.numeric(Freq), breaks = c(-1, 1.1, 5.1, 10.1, 
                                   25.1, 50.1, 100.1, 250.1, 500.1, 1000.1, 2000.1, max(Freq, na.rm = TRUE)+1),
                     labels = c("0-1", "1-5", "5-10", "10-25", "25-50",
                                "50-100", "100-250", "250-500", "500-1000", "1000-2000", ">2000"))) %>% 
  mutate(count = factor(as.character(count), levels = rev(levels(count)))) %>% 
  
  ggplot(mapping = aes(x = cluster_id, y = sample_id, fill = count)) +
    geom_tile(colour = "white", linewidth = 0.3) +
  
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Sample"))+
    labs(x = "Phenotype", y = " sample_id", title = "Phenotype-Sample Distribution for mouse_id = c3m2")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +
    
    # color palette
    #scale_fill_manual(values = rev(brewer.pal(11, "RdGy")), na.value = "azure4") + #RdGy RdBu
    scale_fill_manual(values = brewer.pal(11, "RdBu"), na.value = "azure4") + 
  
    # theme
    theme_grey(base_size = 7)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.2, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     )+
    #scale_x_discrete(guide = guide_axis(n.dodge = 2)) # separate x label into two levels
    scale_x_discrete(guide = guide_axis(angle = 45))

```

# phenotype-sample distribution heatmap for c3m5
```{r}
p_c3m5 <- MB_Xeno_c3m5%>% 
   with(table(sample_id, cluster_id)) %>% 
  as.data.frame() %>% 
  mutate(sample_id = paste("Sample", sample_id)) %>%
  mutate(count = cut(as.numeric(Freq), breaks = c(-1, 1.1, 5.1, 10.1, 
                                   25.1, 50.1, 100.1, 250.1, 500.1, 1000.1, 2000.1, max(Freq, na.rm = TRUE)+1),
                     labels = c("0-1", "1-5", "5-10", "10-25", "25-50",
                                "50-100", "100-250", "250-500", "500-1000", "1000-2000", ">2000"))) %>% 
  mutate(count = factor(as.character(count), levels = rev(levels(count)))) %>% 
  
  ggplot(mapping = aes(x = cluster_id, y = sample_id, fill = count)) +
    geom_tile(colour = "white", linewidth = 0.3) +
  
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Sample"))+
    labs(x = "Phenotype", y = " sample_id", title = "Phenotype-Sample Distribution for mouse_id = c3m5")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +
    
    # color palette
    #scale_fill_manual(values = rev(brewer.pal(11, "RdGy")), na.value = "azure4") + #RdGy RdBu
    scale_fill_manual(values = brewer.pal(11, "RdBu"), na.value = "azure4") + 
  
    # theme
    theme_grey(base_size = 7)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.2, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     )+
    #scale_x_discrete(guide = guide_axis(n.dodge = 2)) # separate x label into two levels
    scale_x_discrete(guide = guide_axis(angle = 45))

```

# phenotype-sample distribution heatmap for c5m1
```{r}
p_c5m1 <- MB_Xeno_c5m1%>% 
   with(table(sample_id, cluster_id)) %>% 
  as.data.frame() %>% 
  mutate(sample_id = paste("Sample", sample_id)) %>%
  mutate(count = cut(as.numeric(Freq), breaks = c(-1, 1.1, 5.1, 10.1, 
                                   25.1, 50.1, 100.1, 250.1, 500.1, 1000.1, 2000.1, max(Freq, na.rm = TRUE)+1),
                     labels = c("0-1", "1-5", "5-10", "10-25", "25-50",
                                "50-100", "100-250", "250-500", "500-1000", "1000-2000", ">2000"))) %>% 
  mutate(count = factor(as.character(count), levels = rev(levels(count)))) %>% 
  
  ggplot(mapping = aes(x = cluster_id, y = sample_id, fill = count)) +
    geom_tile(colour = "white", linewidth = 0.3) +
  
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Sample"))+
    labs(x = "Phenotype", y = " sample_id", title = "Phenotype-Sample Distribution for mouse_id = c5m1")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +
    
    # color palette
    #scale_fill_manual(values = rev(brewer.pal(11, "RdGy")), na.value = "azure4") + #RdGy RdBu
    scale_fill_manual(values = brewer.pal(11, "RdBu"), na.value = "azure4") + 
  
    # theme
    theme_grey(base_size = 7)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.2, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     )+
    #scale_x_discrete(guide = guide_axis(n.dodge = 2)) # separate x label into two levels
    scale_x_discrete(guide = guide_axis(angle = 45))
#  ggsave("Phenotype_Sampling.jpeg", path = "/Users/henzhwang/Desktop/", width = 10, height = 7.5)
```

```{r}
p_c1m1
p_c1m4 
p_c2m1 
p_c2m2 
p_c3m2 
p_c3m5 
p_c5m1 
```

```{r}
# number of cell_type for each cluster_id in each treatment 
# percentage of each treatment in each cluster_id
MB_Xeno %>% 
  group_by(cluster_id) %>% 
  summarise(cluster_n = n(),
            n_ct = n_distinct(treatment_id), 
            AraC_D2 = as.character(as.list(unique(treatment_id))[1]),
            AraC_D2_n = sum(treatment_id == as.character(as.list(unique(treatment_id))[1])),
            ct = as.character(as.list(unique(treatment_id))[2]),
            ct_n = sum(treatment_id == as.character(as.list(unique(treatment_id))[2])),
            AraC_D9 = as.character(as.list(unique(treatment_id))[3]),
            AraC_D9_n = sum(treatment_id==as.character(as.list(unique(treatment_id))[3]))) %>% 
  mutate(AraC_D2_f = formattable::percent(AraC_D2_n / cluster_n),
         ct_f = formattable::percent(ct_n / cluster_n),
         AraC_D9_f = formattable::percent(AraC_D9_n / cluster_n)) %>% 
  dplyr::relocate(AraC_D2_f, .after = AraC_D2_n) %>% 
  dplyr::relocate(ct_f, .after = ct_n) %>% 
  dplyr::relocate(AraC_D9_f, .after = AraC_D9_n) %>% 
  #knitr::kable()
  formattable::formattable(list(AraC_D2_f = color_bar(color = "lightblue"),
                           ct_f = color_bar(color = "lightpink"),
                           AraC_D9_f = color_bar(color = "lightgreen"))) # for html


```


```{r}
as.data.frame(with(MB_Xeno, table(cluster_id, treatment_id))) %>%
  ggplot() +
    #geom_tile(aes(x = cluster_id, y = mm, fill = Freq)) +
    geom_tile(aes(x = treatment_id, y = cluster_id, fill = Freq)) +
  
    # colour palettes
    scale_fill_distiller(palette = "RdBu", na.value = "azure4")+
    # labels
    guides(fill=guide_legend(title = "Number of Cells \nin Each Cluster"))+
    labs(x = "Phenotype", y = " cluster_id", 
         title = "Phenotype-Cluster Distribution MB_Xeno")+
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_discrete(expand = c(0, 0)) +

    # theme
    theme_grey(base_size = 10)+
    theme(legend.position = "right", legend.direction = "vertical",
        legend.title = element_text(colour = "black", size = 6),
        legend.margin = margin(grid::unit(0, "cm")),
        legend.text = element_text(colour = "grey40", size = 6, face = "bold"),
        legend.key.height = grid::unit(0.8, "cm"),
        legend.key.width = grid::unit(0.2, "cm"),
        axis.text.x = element_text(size = 7, colour = "grey40"),
        axis.text.y = element_text(size = 5, vjust = 0.2, colour = "grey40"),
        axis.ticks = element_line(linewidth = 0.4),
        plot.background = element_blank(),
        panel.border = element_blank(),
        plot.margin = margin(0.7, 0.4, 0.1, 0.2, "cm"),
        plot.title = element_text(colour = "grey40", hjust = 0, size = 11, face = "bold")
     ) +
    scale_x_discrete(guide = guide_axis(angle = 45))
```












