---
title: "MB_Xeno Linear Mixed Model"
author: "Xinyue_Cui"
date: "2023-10-07"
output: html_document
---
# packages
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(cytomapper)
library(ggplot2)
library(formattable)
library(magrittr)
library(lme4)
library(car)
BiocManager::install("phyloseq")
Yes
library(ggspavis)
```

# The dataset MB_ Xeno
```{r}
MB_Xeno <- read_csv("/Users/cuixinyue/Desktop/MB_Xeno/MIBI-MB-Xeno-Pilot_scdata_Spatial.csv")

# colnames(MB_Xeno)
# Now, I want to create a new column called tissue_id, which is the unique combination of experience_id and mouse_id. We get 9 unique tissue in total.
MB_Xeno$tissue_id <- paste(MB_Xeno$mouse_id, MB_Xeno$experiment_id, sep="_")

# Convert tissue_id to a factor based on the order
custom_order <- c("c1m1_non-engrafted","c3m2_CB","c5m1_AML", "c2m2_non-engrafted","c1m4_CB","c2m1_AML", "c3m5_non-engrafted", "c2m1_CB","c2m2_AML")
MB_Xeno$tissue_id <- factor(MB_Xeno$tissue_id, levels = custom_order)
# Convert treatment_id to a factor based on the order
custom_order_1 <- c("Ct","AraC_D2","AraC_D9") 
MB_Xeno$treatment_id <- factor(MB_Xeno$treatment_id,levels = custom_order_1)


MB_Xeno$treatment_id <- as.factor(as.character(MB_Xeno$treatment_id))
MB_Xeno$mouse_id <- as.factor(as.character(MB_Xeno$mouse_id))
MB_Xeno$tissue_id <- as.factor(as.character(MB_Xeno$tissue_id))
MB_Xeno$sample_id <- as.factor(as.character(MB_Xeno$sample_id))
MB_Xeno$experiment_id <- as.factor(as.character(MB_Xeno$experiment_id))

table(MB_Xeno$tissue_id)
dim(MB_Xeno)
table(MB_Xeno$treatment_id)
```

create a contingency table that displays the frequency counts of the combinations between the tissue_id and treatment_id variables from the MB_Xeno dataset
```{r}
table(MB_Xeno$tissue_id,MB_Xeno$treatment_id)
```
generate contingency tables to examine the relationship between the 'tissue_id' and three other variables: 'BlockID_1', 'BlockID_2', and 'FOV', within the 'MB_Xeno' dataset
```{r}
# get a distribution plot of CD45_asinh
ggplot(MB_Xeno, aes(x = CD45_asinh)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of CD45_asinh",
       x = "Variable Value",
       y = "Density")

ggplot(MB_Xeno, aes(x =treatment_id, y=CD45_asinh))+geom_violin()
```
# log transformation of all protein markers
```{r}
#Logistic transformation of marker
MB_Xeno$logit.CD45_asinh <- log(MB_Xeno$CD45_asinh)
# View the subset data
head(MB_Xeno)

# Convert factor or character column to numeric
MB_Xeno$logit.CD45_asinh <- as.numeric(as.character(MB_Xeno$logit.CD45_asinh))
# Check for NAs or infinite values
sum(is.na(MB_Xeno$logit.CD45_asinh))
sum(is.infinite(MB_Xeno$logit.CD45_asinh))
MB_Xeno$logit.CD45_asinh[is.infinite(MB_Xeno$logit.CD45_asinh)] <- -9
min(MB_Xeno$logit.CD45_asinh)
max(MB_Xeno$logit.CD45_asinh)
# Remove rows with infinite values in the  MB_Xeno$logit.CD45_asinh column
MB_Xeno <- MB_Xeno[!is.infinite(MB_Xeno$logit.CD45_asinh), ]
# get a distribution plot of log.dsDNA
ggplot(MB_Xeno, aes(x =  MB_Xeno$logit.CD45_asinh)) +
  geom_density(fill = "blue", alpha = 0.5)  +
  xlim(-10,2)+
  labs(title = "Density Plot of logistic transformation of marker ",
       x = "Variable Value",
       y = "Density")
```
Interpret of LMM
Intercept :the estimated mean value for the group of  Ct treatment. 
Treatement_idAraC_D2: the difference in the response between AraC_D2 and Ct.
Treatement_idAraC_D9: the difference in the response between AraC_D9 and Ct.

```{r}
head(MB_Xeno)
levels(MB_Xeno$treatment_id)
MB_Xeno.lme2 <- lmer(logit.CD45_asinh ~ treatment_id + (treatment_id| tissue_id ), data= MB_Xeno)
summary(MB_Xeno.lme2 )
ranef(MB_Xeno.lme2)
fixef(MB_Xeno.lme2)
```
# Build a data set for graph
```{r}
# Selecting the desired columns
subset_MB_Xeno <- MB_Xeno %>%
  select(sample_id, experiment_id, treatment_id,tissue_id, logit.CD45_asinh)
subset_MB_Xeno$type <- rep(c("observations"), each = length(subset_MB_Xeno$sample_id))
head(subset_MB_Xeno)
summary(subset_MB_Xeno)

# Convert the class of treatment_id and tissue_id column to factor
str(subset_MB_Xeno)
custom_order <- c("c1m1_non-engrafted","c3m2_CB","c5m1_AML", "c2m2_non-engrafted","c1m4_CB","c2m1_AML", "c3m5_non-engrafted", "c2m1_CB","c2m2_AML")
subset_MB_Xeno$tissue_id <- factor(MB_Xeno$tissue_id, levels = custom_order)
# Convert treatment_id to a factor based on the order
custom_order_1 <- c("Ct","AraC_D2","AraC_D9") 
subset_MB_Xeno$treatment_id <- factor(MB_Xeno$treatment_id,levels = custom_order_1)
# LMM on subset_MB_Xeno

df <- data.frame(ranef(MB_Xeno.lme2))
# Ct data
slope_Ct <- data.frame(
  tissue_id = df$grp[df$term == "(Intercept)"],
  logit.CD45_asinh = df$condval[df$term == "(Intercept)"]
)
  slope_Ct$treatment_id <- rep(c("Ct"), each = 9)
  slope_Ct$experiment_id <- rep(c("non-engrafted","CB","AML"),each = 3)
  slope_Ct$sample_id  <- NA
# AraC_D2 data
  slope_AraC_D2  <- data.frame(
  tissue_id = df$grp[df$term == "treatment_idAraC_D2"],
  logit.CD45_asinh =df$condval[df$term == "(Intercept)"]+ df$condval[df$term == "treatment_idAraC_D2"]
)
  slope_AraC_D2$treatment_id <- rep(c("AraC_D2"), each = 9)
  slope_AraC_D2$experiment_id <- rep(c("non-engrafted","CB","AML"),each = 3)
  slope_AraC_D2$sample_id  <- NA
# AraC_D9 data
  slope_AraC_D9 <- data.frame(
  tissue_id = df$grp[df$term == "treatment_idAraC_D9"],
  logit.CD45_asinh = df$condval[df$term == "(Intercept)"]+df$condval[df$term == "treatment_idAraC_D9"]
)
  slope_AraC_D9$treatment_id <- rep(c("AraC_D9"), each = 9)
  slope_AraC_D9$experiment_id <- rep(c("non-engrafted","CB","AML"),each = 3)
  slope_AraC_D9$sample_id  <- NA
# Take a look at the data
slope <- rbind(slope_Ct, slope_AraC_D2, slope_AraC_D9)
slope$type <- rep(c("Mixed model"))
# add a row about fixed
# data.frame(fixef(MB_Xeno.lme1))
fixed <- data.frame(
  logit.CD45_asinh = rep(c(-1.6086496	,-1.6086496-0.3015021	,-1.6086496-1.1527743),each = 9),
  experiment_id = NA ,
  sample_id = NA,
  treatment_id = rep(c("Ct","AraC_D2","AraC_D9"), each = 9),
  tissue_id = rep(c("c1m1_non-engrafted","c3m2_CB","c5m1_AML", "c2m2_non-engrafted","c1m4_CB","c2m1_AML", "c3m5_non-engrafted", "c2m1_CB","c2m2_AML"), times = 3),
  type = "all"
)
# Combine the data
plot_data <-rbind(subset_MB_Xeno, slope,fixed)
plot_data
```

```{r}
custom_order <- c("c1m1_non-engrafted","c3m2_CB","c5m1_AML", "c2m2_non-engrafted","c1m4_CB","c2m1_AML", "c3m5_non-engrafted", "c2m1_CB","c2m2_AML")
plot_data$tissue_id <- factor(plot_data$tissue_id, levels = custom_order)
# Convert treatment_id to a factor based on the order
custom_order_1 <- c("Ct","AraC_D2","AraC_D9") 
plot_data$treatment_id <- factor(plot_data$treatment_id,levels = custom_order_1)

#plot
ggplot(plot_data ,aes(x=treatment_id,y=logit.CD45_asinh,color = type))+geom_point()+facet_wrap(~tissue_id, ncol = 3, strip.position = "top")+
  theme_minimal() 
```


```{r}
df <- data.frame(ranef(MB_Xeno.lme1))
df$condval
intercept_condvals <- df$condval[df$term == "(Intercept)"]
head(df)
ggplot(df,aes(x=term,y=condval))+geom_point()+facet_wrap(~grp, ncol = 3, strip.position = "top")
ggplot(df,aes(x = term, y=condval,color = grp))+geom_point()
ggplot(df,aes(x = grp, y=condval,color = term))+geom_point()
```
Slope data set
```{r}
intercept_condvals <- df$condval[df$term == "(Intercept)"]
treatment_idAraC_D2_condvals <- df$condval[df$term == "treatment_idAraC_D2"]
treatment_idAraC_D9_condvals <- df$condval[df$term == "treatment_idAraC_D9"]
slope <- data.frame(
  tissue = df$grp[df$term == "(Intercept)"],
  intercept = df$condval[df$term == "(Intercept)"],
  treatment_idAraC_D2 = df$condval[df$term == "treatment_idAraC_D2"],
  treatment_idAraC_D9 = df$condval[df$term == "treatment_idAraC_D9"]
)
ggplot(slope,aes(x = tissue, y=intercept))+geom_point()
```




```{r}
table(MB_Xeno$treatment_id)
table(MB_Xeno$tissue_id)

ggplot(MB_Xeno, aes(x=treatment_id, y=logit.dsDNA_asinh_znorm)) +ylim(-2,4)+
  geom_violin(trim=TRUE) +
  geom_smooth(method="lm", se=FALSE) +
  stat_summary(fun=mean, geom="point", shape=10, size=1, color="red") + # Adding average point
  facet_wrap(~tissue_id, ncol = 3, strip.position = "top") + 
  theme_minimal() +
  labs(title="Pairwise plot of treatment_id and Average logit.dsDNA_asinh_znorm for each tissue_id")
```
```{r}
# Generate data frame for the straight line
line_df <- data.frame(
  treatment_id = rep(c(1,2,3), each = 3),
  
  logit.dsDNA_asinh_znorm =  + 10.46729  * rep(0:9, each = 1),
  Subject = rep(unique(sleepstudy$Subject), each = length(0:9))
)
# Plot
ggplot(sleepstudy, aes(x=Days, y=Reaction)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) +
  geom_line(data = line_data, aes(x = Days, y = Reaction, group = Subject), color = "red") +  # Adding line with slope of 2
  facet_wrap(~Subject, ncol = 6, strip.position = "top") + 
  theme_minimal()
```

```{r}
# Plot
ggplot(MB_Xeno, aes(x = treatment_id, y = logit.dsDNA_asinh_znorm, group = 1)) +
  geom_line() +
  geom_point() +   # Optional, to display points as well
  theme_minimal()
```

```{r}
# Convert treatment_id to numeric
MB_Xeno$treatment_num <- as.numeric(as.factor(MB_Xeno$treatment_id))
table(MB_Xeno$treatment_num)
# Determine the intercept and slope
intercept <- 5  # Example value, adjust as needed
slope <- 0.5   # Example value, adjust as needed

ggplot(MB_Xeno, aes(x = treatment_num, y = logit.dsDNA_asinh_znorm)) +
  geom_line(aes(group = 1)) +
  geom_point() +
  # geom_abline(intercept = intercept, slope = slope, linetype = "dashed", color = "blue") +  # Adds a line with predetermined slope
  scale_x_continuous(breaks = MB_Xeno$treatment_num, labels = MB_Xeno$treatment_id) + 
  facet_wrap(~ _id, ncol = 3, strip.position = "top") + # Ensures categorical labels are shown on x-axis
  theme_minimal()
```
```{r}
colnames(MB_Xeno)
MB_Xeno.lme2 <- lmer(logit.dsDNA_asinh_znorm ~ treatment_num + (treatment_num| tissue_id ), data= MB_Xeno)
ranef(MB_Xeno.lme2)
fixef(MB_Xeno.lme2)
```


X be the treatment id
```{r}
ggplot(MB_Xeno, aes(x=treatment_num, y=logit.dsDNA_asinh_znorm)) +geom_line()+
  geom_point() +
  facet_wrap(~tissue_id, ncol = 3, strip.position = "top") + 
  scale_y_continuous(limits=c()) +
  theme_minimal() +
  labs(title="Pairwise plot of treatment_id and Average logit.dsDNA_asinh_znorm for each tissue_id") 
```
