---
title: "MB_Xeno Linear Mixed Model"
author: "Xinyue_Cui"
date: "2023-10-07"
output: html_document
---
# packages
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(cytomapper)
library(ggplot2)
library(formattable)
library(magrittr)
library(lme4)
library(car)
library(Metrics)
library(performance)
library(lattice)
library(caret)
```
# The dataset MB_ Xeno
```{r}
MB_Xeno <- read_csv("/Users/cuixinyue/Desktop/MB_Xeno/MIBI-MB-Xeno-Pilot_scdata_Spatial.csv")
head(MB_Xeno)
colnames(MB_Xeno)
# colnames(MB_Xeno)
# Now, I want to create a new column called tissue_id, which is the unique combination of experience_id and mouse_id. We get 9 unique tissue in total.
MB_Xeno$tissue_id <- paste(MB_Xeno$mouse_id, MB_Xeno$experiment_id, sep="_")
table(MB_Xeno$tissue_id)
MB_Xeno$tissue_id <- as.factor(as.character(MB_Xeno$tissue_id))
MB_Xeno$treatment_id <- as.factor(as.character(MB_Xeno$treatment_id))
MB_Xeno$Run <- as.factor(as.character(MB_Xeno$Run))
MB_Xeno$mouse_id <- as.factor(as.character(MB_Xeno$mouse_id))
# Convert the first 12 column of MB_Xeno to factor
# for (i in 1:12) {
#   if (!is.factor(MB_Xeno[[i]])) {
#     MB_Xeno[[i]] <- as.factor(MB_Xeno[[i]])
#   }}
# str(MB_Xeno[, 1:12])

```

```{r}
#Logistic transformation of dsDNA_asinh_znorm
MB_Xeno$logit.dsDNA_asinh_znorm <- log(MB_Xeno$dsDNA_asinh_znorm/(1-MB_Xeno$dsDNA_asinh_znorm))
# View the subset data

# Convert factor or character column to numeric
MB_Xeno$logit.dsDNA_asinh_znorm <- as.numeric(as.character(MB_Xeno$logit.dsDNA_asinh_znorm))
# Check for NAs or infinite values
sum(is.na(MB_Xeno$logit.dsDNA_asinh_znorm))
sum(is.infinite(MB_Xeno$logit.dsDNA_asinh_znorm))
min(MB_Xeno$logit.dsDNA_asinh_znorm)
max(MB_Xeno$logit.dsDNA_asinh_znorm)
# Remove rows with infinite values in the logit.dsDNA_asinh_znorm column
MB_Xeno <- MB_Xeno[!is.infinite(MB_Xeno$logit.dsDNA_asinh_znorm), ]
# get a distribution plot of log.dsDNA
ggplot(MB_Xeno, aes(x = logit.dsDNA_asinh_znorm)) +
  geom_density(fill = "blue", alpha = 0.5)  +
  labs(title = "Density Plot of logistic transformation of dsDNA_asinh_znorm ",
       x = "Variable Value",
       y = "Density")
```


#1. Train Test split of all observation 
(I am trying to avoid splitting data from the same tissue between training and testing sets to prevent data leakage. )
##1.1 Split the train test
70% of the observations is about 393558, So I put tissue_id of c1m1_non-engrafted,c1m4_CB,c2m1_AML, c2m1_CB, c2m2_AML in the train set. （total of 384858 observations). tissue_id in (c2m2_non-engrafted, c3m2_CB, c3m5_non-engrafted, c5m1_AML) into test set
```{r}
tissue_id_select <- c("c1m1_non-engrafted","c1m4_CB","c2m1_AML", "c2m1_CB","c2m2_AML" )

train <- MB_Xeno[MB_Xeno$tissue_id %in% tissue_id_select ,]
test <- MB_Xeno[!MB_Xeno$tissue_id %in% tissue_id_select ,]
```

```{r}
# get a distribution plot of dsDNA_asinh_znorm
ggplot(train, aes(x = dsDNA_asinh_znorm)) +
  geom_density(fill = "blue", alpha = 0.5) +
  xlim(0, 1) +
  labs(title = "Density Plot of dsDNA_asinh_znorm ",
       x = "Variable Value",
       y = "Density")
```
##1.2Model Fitting and Selection
```{r}
train.lme1 <- lmer(logit.dsDNA_asinh_znorm ~ treatment_id + ( treatment_id| tissue_id ), data= train )
summary(train.lme1)
AIC(train.lme1)
BIC(train.lme1)
# table(MB_Xeno$Run, MB_Xeno$tissue_id)
train.lme2 <- lmer(logit.dsDNA_asinh_znorm ~ treatment_id + (treatment_id|Run ), data= train )
summary(train.lme2)
AIC(train.lme2)
BIC(train.lme2)
```
## 1.3 predict on test set
```{r}
# Predict on the testing set (hat(y*))
pred_dsDNA <- predict(train.lme1, newdata = test, re.form = NA)
hist(x=pred_dsDNA )
length(pred_dsDNA )
# Actual (y*)
actual_dsDNA<- test$logit.dsDNA_asinh_znorm
length(actual_dsDNA )
# Calculate RMSE
sqrt(mean((test$logit.dsDNA_asinh_znorm - pred_dsDNA )^2))

print(paste("RMSE:", rmse_value))
# Calculate MAE
mae_value <- mae(actual_dsDNA, pred_dsDNA)
print(paste("MAE:", mae_value))

# Calculate R-squared
r_squared <- r2(train.lme1)
print(paste("R-squared:", r_squared))
```
##1.4 Check residuals
```{r}
plot(train.lme1)
hist(residuals(train.lme1))
qqnorm(residuals(train.lme1))
qqline(residuals(train.lme1))
```
## 1.5 Visualizing random effects
```{r}
ranef_plot <- dotplot(ranef(train.lme1, condVar = TRUE))
print(ranef_plot)
```



# 2. Split for training and test data according to stations
## 2.1 Split
```{r}
sample_train <- sample(unique(MB_Xeno$sample_id), size = length(unique(MB_Xeno$sample_id))*0.7)
training <- MB_Xeno[MB_Xeno$sample_id %in% sample_train,]
testing <- MB_Xeno[!MB_Xeno$sample_id %in% sample_train,]
colnames(MB_Xeno)
```
## 2.2 fitting a LMM
```{r}
model_lmm1 <- lmer(logit.dsDNA_asinh_znorm ~  treatment_id + (treatment_id|tissue_id), data = training)
summary(model_lmm1)
lmm1_prd_1 <- predict(model_lmm1, newdata = testing)
mean((testing$logit.dsDNA_asinh_znorm - lmm1_prd_1 )^2)
```
