---
title: "05_MB_Xeno_LDA_multiChains"
knit: (function(input, encoding) {
  rmarkdown::render(input = input,
                    output_dir = here::here(),
                     knit_root_dir = getwd())})
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document: 
    toc: yes
date: "2023-02-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "../")
#knitr::opts_knit$set(output.dir = "../Output_Documents/")
options(dplyr.summarise.inform = FALSE)
```


## Library
```{r library, warning=FALSE, message=FALSE}
library(rstan)
library(dplyr)
library(ggplot2)
library(randomcoloR)
library(SpatialExperiment)
library(here)
```


```{r plot theme}
theme_set(theme_minimal())
theme_update(
  text = element_text(size = 10),
  legend.text = element_text(size = 10)
)


# plot theme
remove_x <- theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
)

remove_y <- theme(
  axis.text.y = element_blank(),
  axis.ticks.y = element_blank(),
  axis.title.y = element_blank()
)



# force not to use scientific notation
options(scipen=999)
```

```{r}
library(here)
set_here("/Users/cuixinyue/Desktop/MB_Xeno")
```

## Dataset

Load the TNBC SPE object.
```{r dataset, warning=FALSE, message=FALSE}
load(file = here("05_MB_Xeno_DA_multiChains","02_MB_Xeno_spe.rds"))
spe
```

## Pre-process
### Document-Term Matrix

We want to create the Document-Term matrix of the dataset where the rows are the sample identity and columns are the cell types fill with counts. The expected dimension of the DTM is 9 rows by 33 columns.

```{r dtm}
dtm <- table(spe$tissue_id, spe$cluster_id)
dim_names <- dimnames(dtm)
dim(dtm)
```
### Posterior Sampling

We want to estimate the parameters $\theta$ and $B = \left[\beta_{1}, \cdots, \beta_{K} \right]^{T}$, and we estimate using Hamiltonian Monte Carlo(HMC) No-U-Turn-Sampler(NUTS) with one chain and 500, 1000, and 2000 iterations. Out of these iterations, half of the iteration were used as warm-up samples.

We will apply `get_lda_result()` to perform LDA and save results.

```{r lda result, eval=FALSE}
source(here::here("06_LDA_scripts", "06_get_lda_result.R"))

samples <- get_lda_result(K = 3, alpha = 0.8, gamma = 0.8,
                          dtm = dtm, iter = 500, chains = 2,
                          file_fold = "07_LDA_multiChains", save_file = TRUE)

```







