---
title: "Diagnostic_of _LDA"
author: "Xinyue_Cui"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
#library(phyloseq)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)
#library(genefilter) 
library(rstan)
library(randomcoloR)
library(SpatialExperiment)
#library(DESeq2)
# library(here)
``` 

```{r}
load(file = here::here("Output","RData","02_SPE","02_MB_Xeno_spe.rds"))
spe
```

#Trace plot
```{r}

  # setting seed will align topics
  # not applying seed, topics will differ but similar probabilities
  set.seed(19991116)
  
  K <- 3
  alpha <- 0.8
  gamma <- 0.8
  x <- dtm
  iter <-  500
  chains <- 2
  # View(x)
  dimnames(x) <- NULL
  
  
  # tissues <- get_lda_result(K = 5, alpha = 0.8, gamma = 0.8,dtm = dtm, iter = 2000, chains = 2, file_fold = "05_LDA_multiChains", save_file = TRUE)
  # theta[d] ~ dirichlet(alpha), alpha pseudo-count for each topic
  # beta[k] ~ dirichlet(gamma), gamma pseudo-count for each ASV in each topic
  
  # n is DTM
  stan.data <- list(
    K = K,
    V = ncol(x),
    D = nrow(x),
    n = x,
    alpha = rep(alpha, K),
    gamma = rep(gamma, ncol(x))
  )
  
  # file name
  fileN <- paste0("JABES_Endo_all_K_",
                  K,
                  "_ite_",
                  iter,
                  ".RData")
  

  
  # apply dataset
  t1 <- proc.time()
  stan.fit_1 <- stan(
    file = here::here("Notebooks", "06_LDA_scripts", "06_lda.stan"),
    data = stan.data,
    # must be list object
    iter = 500,
    chains = 2,
    sample_file = NULL,
    diagnostic_file = NULL,
    cores = 4,
    control = list(adapt_delta = 0.9),
    save_dso = TRUE,
    algorithm = "NUTS"
  )
  proc_time <- proc.time() - t1   # processing time
  # save file with specified name
  
traceplot(stan.fit_1)
dim(stan.fit_1)

  # save(stan.fit,  file = here::here("Output", "RData","04_MB_Xeno_LDA", paste0(fileN)))
  
 # res <- rstan::extract(
  #   stan.fit,
  #   permuted = TRUE,
  #   inc_warmup = FALSE,
  #   include = TRUE
  # )
  # 

```
```{r}
e <- extract(stan.fit_1, permuted = FALSE) # return a list of arrays
str

length(e[1,1,])
dim(e[,2,])

dim(e[,1,])
dim(e[,2,])
```


```{r}
stan.fit_1
dim(stan.fit_1)
# Assuming 'fit' is your stan.fit object
theta_samples <- extract(stan.fit_1)$theta
dim(extract(stan.fit_1)$theta)
# 'theta_samples' will be an array or list (depending on the parameter dimensions) 

# Assuming 'fit' is your stan.fit object and you want to include warm-up iterations
theta_with_warmup <- as.array(stan.fit_1, pars = "theta", inc_warmup = TRUE)

# Now, 'theta_with_warmup' includes warm-up iterations for 'theta'


??traceplot
traceplot(stan.fit_1,pars = c("theta"), inc_warmup = FALSE, nrow = 9)
```

```{r}
# install.packages("loo")
library(loo)
```

# Align matrix
## Dataset

Load the MB_Xeno SPE object.
```{r dataset, warning=FALSE, message=FALSE}
load(file = here("Output","RData","02_SPE","02_MB_Xeno_spe.rds"))
spe
```

## Pre-process
### Document-Term Matrix

We want to create the Document-Term matrix of the data set where the rows are the sample identity and columns are the cell types fill with counts. The expected dimension of the DTM is 9 rows by 33 columns.

```{r dtm}
dtm <- as.data.frame.matrix(table(spe$tissue_id, spe$cluster_id)) %>% as.matrix()
dtm <- as.matrix(table(spe$tissue_id, spe$cluster_id))
dim(dtm)
dim_names <- dimnames(dtm)
col_names <- colnames(dtm)
```
### Posterior Sampling

We want to estimate the parameters $\theta$ and $B = \left[\beta_{1}, \cdots, \beta_{K} \right]^{T}$, and we estimate using Hamiltonian Monte Carlo(HMC) No-U-Turn-Sampler(NUTS) with one chain and 500, 1000, and 2000 iterations. Out of these iterations, half of the iteration were used as warm-up samples.

We will apply `get_lda_result()` to perform LDA and save results.

```{r lda result, eval=FALSE}
source(here::here("Notebooks","06_LDA_scripts", "06_get_lda_result.R"))

tissues <- get_lda_result(K = 3, alpha = 0.8, gamma = 0.8,
                          dtm = dtm, iter = 500, chains = 2,
                          file_fold = "05_LDA_multiChains", save_file = TRUE)
```
Load the existing LDA estimations.
```{r}
# source(here::here("Notebooks",
#                   "06_LDA_scripts", "06_load_lda_result.R"))
# 
# file_fold = "05_LDA_multiChains"
# lda_file = "JABES_Endo_all_K_3_ite_500.RData"
# 
# tissues <- load_lda_result(
#   file = here::here("Output", "RData",
#                    file_fold, lda_file)
# )
```

```{r}
K = 3
iter = 500
chain = 2
tissues_1 <- tissues[[1]]
tissues_2 <- tissues[[2]]
dim(stan.fit)
```

```{r}
# dim(tissues_1$theta)
dim(tissues_2$theta)

# theta estimation
theta <- tissues_2$theta
dim(theta)

# beta estimation
beta <- tissues_2$beta
dim(beta)

```

### Alignment matrix
```{r}
library(abind)
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_alignmentMatrix.R"))

aligned_matrix <- alignmentMatrix(
  theta = theta,
  spe = spe,
  K = K,
  iter = iter,
  chain = chain,
  #cor_method = "cosine",
  TissueID_name = "tissue_id"
)
```

### Align topic proportion

We align the topic proportion in each specimen across two chains.
```{r align-theta}
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_thetaAligned.R"))

theta_aligned <- thetaAligned(
  theta = theta, 
  K = K, 
  aligned = aligned_matrix, 
  iter = iter,
  chain = chain,
  )
# theta_aligned 
dim(theta_aligned)
theta_aligned[,1,1]
```

```{r align-beta}
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_betaAligned.R"))

beta_aligned <- betaAligned(
  beta = beta, 
  K = K, 
  aligned = aligned_matrix, 
  iter = iter,
  chain = chain,
  ) 
# beta_aligned 
```
#Traceplot after alignment
```{r}
# Load necessary libraries
library(ggplot2)

# Your theta_aligned matrix
# theta_aligned <- your_data_here

# Split the data into two chains
chain1 <- theta_aligned[1:250, , ]
chain2 <- theta_aligned[251:500, , ]

# Initialize an empty list to store plots
plots <- list()

# Generate plots for each tissue and topic combination
for (tissue in 1:9) {
    for (topic in 1:3) {
        # Extract data for this tissue-topic combination from both chains
        data_chain1 <- data.frame(Iteration = 1:250, Theta = chain1[, tissue, topic], Chain = "Chain 1")
        data_chain2 <- data.frame(Iteration = 1:250, Theta = chain2[, tissue, topic], Chain = "Chain 2")
        data <- rbind(data_chain1, data_chain2)
        
        # Create the plot
        p <- ggplot(data, aes(x = Iteration, y = Theta, color = Chain)) +
            geom_line() +
            labs(title = paste("Tissue", tissue, "Topic", topic),
                 x = "Iteration",
                 y = "Theta Value") +
            theme_minimal() +
            scale_color_manual(values = c("Chain 1" = "blue", "Chain 2" = "red"))
        
        # Store the plot in the list
        plots[[paste("Tissue", tissue, "Topic", topic)]] <- p
    }
}

# Display the plots
# You can access each plot individually, for example, plots[["Tissue 1 Topic 1"]] to view the first plot
# If you want to display all plots in a loop, you can use:
for (plot_name in names(plots)) {
    print(plots[[plot_name]])
}

```

```{r}
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_traceplot_aligned.R"))

traceplot_aligned(theta_aligned, 
                  num_chains = 2 , 
                  num_iterations_per_chain = 250)

```

## Diagnostic after aligned
```{r}
Rhat_theta <- matrix(
    nrow = dim(theta_aligned)[2],
    ncol = dim(theta_aligned)[3]
  )

```
```{r}
#dim(theta_aligned)
#beta_aligned
# Initialize the x_sim array with the desired dimensions
x_sim <- array(NA, dim = c(500, 9, 33))

# Loop over each iteration
for (i in 1:500) {
    # Perform matrix multiplication for the i-th iteration
    # theta_aligned[i,,] selects the i-th 9x3 matrix
    # beta_aligned[i,,] selects the i-th 3x33 matrix
    # The result is a 9x33 matrix for this iteration
    x_sim[i,,] <- theta_aligned[i,,] %*% beta_aligned[i,,]
}

# x_sim now contains the product for each iteration, with dimensions 500x9x33
dim(x_sim)
Rhat(x_sim[,,])
is.matrix(x_sim)
dim(sims)
```

## Visualization

### Plot topic proportion

Plot the topic proportion in specimens and we can save it for publication purposes. We can draw an informative summary of bacterial communities in each phenotype.
```{r plot-topic-prop, fig.width=20, fig.height=8}
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_plotTopicProportion.R"))

p_topicProp <- plotTopicProportion(
  spe = spe,
  theta_aligned = theta_aligned,
  K = K,
  iter = iter,
  chain = chain,
  TissueID_name = "tissue_id"
)
p_topicProp + theme(
  axis.title = element_text(size = 14),
  axis.text = element_text(size = 14),
  plot.title = element_text(size = 14))

#ggsave(filename = "Topic_Prop_5.png", path="/Users/henzhwang/Desktop/",
#       width = 15, height = 10)
```
### Plot ASVs distributions
We plot the ASVs distribution in each topic.
```{r plot cellType Dist, fig.width=10, fig.height=10}
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_plotCellTypeDistribution.R"))

p_cellTypeDist <- plotCellTypeDistribution(
  spe = spe,
  K = K,
  beta_aligned = beta_aligned,
  col_names_beta_hat = c("iterations", "Topic", "Cell.Type", "beta_h")
)

p_cellTypeDist
#ggsave(filename = "cellType_dist_5.png", path="/Users/henzhwang/Desktop/",
#       width = 15, height = 10)
```
## Diagnostic plot
Perform computation of effective sample size for each parameter
```{r}
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_diagnosticsPlot.R"))

p_diag_3_500 <- diagnosticsPlot(
  theta_aligned = theta_aligned,
  beta_aligned = beta_aligned,
  iter = 500,
  chain = 2,
) 
p_diag_3_500
Rhat
```

## Model Asseessement
### model assessment with max 
We want to perform goodness of fit test using the posterior estimates and observed data.
```{r}
# model assessment with max 
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_modelAssessment_max.R"))
#cellTypeIndexToPlot = c(2, 3, 5, 12, 15)

p_model_assessment_max <- modelAssessment_max(dtm = dtm,
                tissues = tissues_2,
                warm_up_iter = NULL,
                iter = 500,
                cellTypeIndexToPlot = c(1:33)
                )
p_model_assessment_max
```

### model assessment with mean 
```{r}
# model assessment with max 
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_modelAssessment_mean.R"))
#cellTypeIndexToPlot = c(2, 3, 5, 12, 15)

p_model_assessment_mean <- modelAssessment_mean(dtm = dtm,
                tissues = tissues_2,
                warm_up_iter = NULL,
                iter = 500,
                cellTypeIndexToPlot = c(1:33)
                )
p_model_assessment_mean
```
### model assessment with mean 
```{r}
# model assessment with max 
source(here::here("Notebooks", 
                  "06_LDA_scripts", "06_modelAssessment_median.R"))
#cellTypeIndexToPlot = c(2, 3, 5, 12, 15)

p_model_assessment_median <- modelAssessment_median(dtm = dtm,
                tissues = tissues_2,
                warm_up_iter = NULL,
                iter = 500,
                cellTypeIndexToPlot = c(1:33)
                )
p_model_assessment_median
```