---
title: "MB_Xeno_LDA_alto"
knit: (function(input, encoding) {
  rmarkdown::render(input = input,
                    output_dir = here::here(),
                     knit_root_dir = getwd())})
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document: 
    toc: yes
date: "2023-11-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "../")
#knitr::opts_knit$set(output.dir = "../Output_Documents/")
options(dplyr.summarise.inform = FALSE)
```

## Library
```{r library, warning=FALSE, message=FALSE}
devtools::install_github("lasy/alto")

library(alto)
library(dplyr)
library(magrittr)
library(ggplot2)
library(tibble)
library(stringr)
library(tidyr)
```

```{r}
dtm
```

```{r}
# ??run_lda_models
```
##Use `alto` on the data
```{r run-lda-models}
lda_varying_params_lists <-  list()
for (k in 1:15) {lda_varying_params_lists[[paste0("k",k)]] <- list(k = k)}

lda_models_1  <- 
  run_lda_models(
    data = dtm,
    lda_varying_params_lists = lda_varying_params_lists
  )

```
We can now align the topics from each consecutive models:
```{r}
aligned_topics_product <- 
  align_topics(
    models = lda_models_1,
    method = "product") 

aligned_topics_transport <- 
  align_topics(
    models = lda_models_1,
    method = "transport") 
```
##Visualize the alignements
```{r alignment-viz}
plot(aligned_topics_product)
plot(aligned_topics_transport)
```

```{r key-number-of-topics}
compute_number_of_paths(aligned_topics_product) %>% 
  plot_number_of_paths() + 
  ggtitle("Method: product")
compute_number_of_paths(aligned_topics_transport) %>% 
  plot_number_of_paths() + 
  ggtitle("Method: transport")
```
```{r plot-with-add-leaves-and-id-numbers}

aligned_topics_transport_12 <- align_topics(lda_models_1[1:12], method = "transport")
plot(aligned_topics_transport_12, add_leaves = TRUE, label_topics = TRUE)

```
We can also evaluate the coherence and ancestry scores of these topics with the `plot` function and the `color_by` option.

## Alignment with diagnostics scores

```{r coherence}
plot(aligned_topics_product, color_by = "coherence")
plot(aligned_topics_transport, color_by = "coherence")
```

Most of the identified topics around $K=12$ are coherent across $K$.


```{r ancestry}
plot(aligned_topics_product, color_by = "refinement")
plot(aligned_topics_transport, color_by = "refinement")
```

Interestingly, by $K=7$, the identified topics correspond to the four Lactobacillus community state types (CST) and to three non-Lactobacillus topics. Among these, one topic, dominated by a specific strain of Gardnerella and Atopobium, remains coherent across models. The two remaining topics at $K=7$ have little overlap and high refinement scores, indicative that these topics successfully identify two distinct groups of communities, which are revealed as robust topics as $K$ is increased. 

The function `plot_beta` allows to visualize the betas for a selection (or all) models.

```{r beta-plot}
plot_beta(aligned_topics_product, models = c("k3", "k7","k12"), threshold = 0.005)
plot_beta(aligned_topics_transport, models = c("k3", "k7","k12"), threshold = 0.005)
```

By $K=18$ topics are sparse and share little overlap between them (Fig \ref{fig:microbiome_figure}d), which may reflect over-fitting.

The `plot_beta` function also has a `color_by` option, which allows to visualize the coherence of topics.

```{r}
plot_beta(aligned_topics_transport, 
          models = c("k3", "k7","k12"), 
          threshold = 0.005, 
          color_by = "coherence")
```

```{r align-topics-viz-figure, echo = FALSE, eval = FALSE}
# 
# figure_full = ggpubr::ggarrange(
#   g_key_topics_product,
#   g_key_topics_transport,
#   g_refinement_product,
#   g_refinement_transport,
#   g_robustness_product, 
#   g_robustness_transport,
#   g_branches_product,
#   g_branches_transport,
#   g_beta_product,
#   g_beta_transport,
#   ncol = 2,
#   nrow = 5,
#   heights = c(1.5,2,2,2,3)
# )
# 
# ggsave(figure_full, filename = "microbiome_figure.png", height = 22, width = 18, units = "cm", scale = 2)
```
```{r}
make_figure <- function(aligned_topics_product,
                        aligned_topics_transport){
  
  
  n_paths_combined <- 
    bind_rows(
      compute_number_of_paths(aligned_topics_product) %>% 
        mutate(method = "product"),
      compute_number_of_paths(aligned_topics_transport) %>% 
        mutate(method = "transport")
    )
  
  
  g_key_topics <- 
    ggplot(n_paths_combined, 
           aes(x = n_topics,
               y = n_paths,
               col = method)) +
    geom_abline(slope = 1, intercept = 0, col = "gray90", size = 2) +
    geom_line() +
    geom_point(size = 1) +
    theme_minimal() +
    xlab("# of topics in model") +
    ylab("# of paths") +
    scale_x_continuous(breaks = 1:max(n_paths_combined$n_topics), minor_breaks = NULL) +
    scale_y_continuous(breaks = 1:max(n_paths_combined$n_topics), minor_breaks = NULL) +
    scale_color_manual("", values = c("coral","blue3")) +
    theme(
      legend.position = "top", 
      #legend.background = element_rect(fill = "white"),
      legend.title = element_blank(),
      legend.box.margin = margin(t = -10, b = -10, unit = "pt")
    )
    topics <- 
    bind_rows(
      aligned_topics_product@topics %>% mutate(method = "product"),
      aligned_topics_transport@topics %>% mutate(method = "transport"),
    ) %>% 
    mutate(method = method %>% factor(., levels = c("transport","product"))) %>% 
    group_by(method, m) %>% mutate(K = n(), r = refinement) %>% ungroup() 
  
  model_summary <- 
    topics %>% 
    group_by(method, K) %>%
    summarize(
      minr = min(r),
      maxr = max(r), 
      r = median(r),
      .groups = "drop"
    )
  g_refinement_distribution <- 
    ggplot(model_summary,
           aes(x = K, y = r)) + 
    geom_hline(yintercept = 0, col = "gray") +
    geom_ribbon(aes(ymin = minr, ymax = maxr), fill = "gray", alpha = 0.5) +
    geom_line(data = model_summary, col = "black", size = 1) +
    geom_point(data = model_summary, col = "black", size = 2) +
    geom_jitter(data = topics, aes(col = path), height = 0, width = 0.05, size = 1) +
    facet_wrap(method ~ ., dir = "v") + 
    scale_x_continuous(breaks = sort(unique(topics$K)), minor_breaks = NULL) +
    scale_y_log10() +
    scale_color_discrete(breaks = sort(unique(model_summary$K))) +
    guides(col = "none")  +
    xlab("# of topics in model") +
    ylab("refinement score")
  
  
  g_coherence_transport <- 
    plot_alignment(
      aligned_topics_transport, 
      color_by = "co", 
      model_name_repair_fun = function(x) str_remove_all(x, "k")
    ) + 
    theme_minimal() +
    theme(legend.position = "right")
  g_paths_transport <- 
    plot_alignment(
      aligned_topics_transport, 
      model_name_repair_fun = function(x) str_remove_all(x, "k")
    ) +
    theme_minimal() 
  
  
  g_beta_transport <- 
    plot_beta(
      aligned_topics_transport, 
      models = c("k3", "k7","k12",last(aligned_topics_transport@topics$m %>% levels())), 
      threshold = 0.003
    )
  figure = 
    ggpubr::ggarrange(
      g_key_topics,
      ggpubr::ggarrange(
        g_paths_transport,
        g_coherence_transport,
        ncol = 2, nrow = 1, 
        widths = c(4,4.5),
        labels = c("b","c")
      ),
      g_refinement_distribution,
      g_beta_transport,
      ncol = 2, nrow = 2,
      heights = c(1, 1.3),
      widths = c(3,8.5),
      labels = c("a","","d","e")
    )
  # ggpubr::ggarrange(
  #   ggpubr::ggarrange(
  #     ggpubr::ggarrange(
  #       g_key_topics,
  #       g_refinement_distribution,
  #       ncol = 1, nrow = 2, 
  #       heights = c(1, 1.2),
  #       labels = "auto"
  #     ),
  #     g_coherence_transport,
  #     g_paths_transport,
  #     ncol = 3, nrow = 1,
  #     widths = c(4,5,4),
  #     labels = c("","b","c")),
  #   g_beta_transport,
  #   ncol = 1,
  #   nrow = 2,
  #   heights = c(1, 1.1),
  #   labels = c("","d")
  # )
  
  figure
}
  
  
```

```{r}
aligned_topics_product_15 <- align_topics(lda_models_1[1:15], method = "product")
aligned_topics_transport_15 <- align_topics(lda_models_1[1:15], method = "transport")

figure_1_15 <- make_figure(aligned_topics_product = aligned_topics_product_15, 
                           aligned_topics_transport = aligned_topics_transport_15)


ggsave(figure_1_15, filename = "figure.png", height = 14, width = 18, units = "cm", scale = 1.5)


```




